Title
======
Practical Machine Learning Assignment

Loading data
=============
```{r}
training <- read.csv('./pml-training.csv', header=T)

testing <- read.csv('./pml-testing.csv', header=T)
```

Data Partitioning
=================
```{r}
library(ggplot2)
library(lattice)
library(caret)
set.seed(127)
training_sample <- createDataPartition(y=training$classe, p=0.7, list=FALSE)
training <- training[training_sample, ]
test <- training[-training_sample, ]
```

Identification of non zero data
===============================
```{r}
all_zero_colnames <- sapply(names(testing), function(x) all(is.na(testing[,x])==TRUE))
pred_data <- names(all_zero_colnames)[all_zero_colnames==FALSE]
pred_data  <- pred_data [-(1:7)]
pred_data  <- pred_data [1:(length(pred_data )-1)]
```

Model building
==============
```{r}
library(caret)
library(rpart)
library(gbm)
library(randomForest)
library(plyr)
fitControl <- trainControl(method='cv', number = 3)
model_cart <- train(
  classe ~ ., 
  data=training[, c('classe', pred_data)],
  trControl=fitControl,
  method='rpart'
)
save(model_cart, file='./ModelFitCART.RData')
model_gbm <- train(
  classe ~ ., 
  data=training[, c('classe', pred_data)],
  trControl=fitControl,
  method='gbm'
)
save(model_gbm, file='./ModelFitGBM.RData')
model_rf <- train(
  classe ~ ., 
  data=training[, c('classe', pred_data)],
  trControl=fitControl,
  method='rf',
  ntree=100
)
save(model_rf, file='./ModelFitRF.RData')
```

Cross validation
================
```{r}
fitControl <- trainControl(method='cv', number = 3)
```

Estimation of out of sample error
=================================
```{r}
predCART <- predict(model_cart, newdata=test)
cmCART <- confusionMatrix(predCART, test$classe)
predGBM <- predict(model_gbm, newdata=test)
cmGBM <- confusionMatrix(predGBM, test$classe)
predRF <- predict(model_rf, newdata=test)
cmRF <- confusionMatrix(predRF, test$classe)
AccuracyResults <- data.frame(
  Model = c('CART', 'GBM', 'RF'),
  Accuracy = rbind(cmCART$overall[1], cmGBM$overall[1], cmRF$overall[1])
)
print(AccuracyResults)
```

Conclusion
==========
Based on the data available,the data provided is able to fit a reasonably sound model with a high degree of accuracy in predicting out of sample observations, which is using the random forest decision tree model (100% accuracy obtained). When fitting a model on all training data columns, some features that are all missing in the validation sample do included non-zero items in the training sample and are used in the decision tree models.